Positivity-Preserving Adaptive Runge--Kutta Methods

### 2020-03-31 latex/overview.tex
references in tables of DOFs and layout

@@

\begin{table}[h!]
\centering    %Generated below============ 
\begin{tabular}{|l c|c c c c c c |} 
 \hline 
Order & &1&2&3&4&5&6 \\ 
 \hline Classical RK4& s=4&3&2&0&0& - & -  \\ 
 SSPRK(10,4)& s=10&9&8&6&4& - & -  \\ 
 Cash-Karp RK5(4)6& s=6&5&4&2&1&0& -  \\ 
 Dormand-Prince RK5(4)7& s=7&6&5&3&1&0& -  \\ 
 \hline 
 \end{tabular}
 \caption{Degrees of freedom for the choice of the weights for some explicit methods.} %Generated above============
 \label{table:DOF_exp}
 \end{table}

 \begin{table}[h!]
\centering   %Generated below============ 
\begin{tabular}{|l c|c c c c c c |} 
 \hline 
Order & &1&2&3&4&5&6 \\ 
 \hline Implicit Euler& s=1&0& - & - & - & - & -  \\ 
 Lobatto~IIIC4& s=4&3&2&1&0&0&0 \\
 Radau~IIA3& s=3&2&1&0&0&0& -  \\
 SDIRK 54& s=5&4&3&1&0& - & -  \\ 
 TR-BDF2& s=3&2&1& - & - & - & -  \\ 
 Im-Euler 2& s=3&2&1& - & - & - & -  \\ 
 Im-Euler 3& s=6&5&4&2& - & - & -  \\ 
 Im-Euler 4& s=10&9&8&6&3& - & -  \\ 
 \hline 
 \end{tabular}
 \caption{Degrees of freedom for the choice of the weights for some implicit methods.} %Generated above============
 \label{table:DOF_imp}
 \end{table}

@@

In Table\,\ref{table:DOF_imp}, the same properties of some implicit
methods are reported.
We can see that the fully implicit methods Lobatto~IIIC4 and Radau~IIA3 require a drastic reduction of the order as expected.
The diagonally implicit SDIRK 54 method only requires an order reduction of one to get one degree of freedom for the weights.
The TR-BDF2 method even allows adaptations without reducing the order.  
The Implicit-Euler extrapolation methods also exhibit degrees of freedom without a reduction of the order.

>>>>

\usepackage{booktabs}
\usepackage{multirow}

@@

\begin{table}[h!]
\centering    %Generated below============ 
% \begin{tabular}{|l c|c c c c c c |}
%  \hline
% Order & &1&2&3&4&5&6 \\
%  \hline Classical RK4& s=4&3&2&0&0& - & -  \\
%  SSPRK(10,4)& s=10&9&8&6&4& - & -  \\
%  Cash-Karp RK5(4)6& s=6&5&4&2&1&0& -  \\
%  Dormand-Prince RK5(4)7& s=7&6&5&3&1&0& -  \\
%  \hline
%  \end{tabular}
  \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lr*6c@{}}
    \toprule
    Method & $s$ & \multicolumn{6}{c}{Order $\tilde p$} \\
    & & 1 & 2 & 3 & 4 & 5 & 6 \\
    \midrule
    Classical RK4 \cite{kutta1901beitrag} & 4 & 3 & 2 & 0 & 0 & --- & --- \\
    SSPRK(10,4) \cite{ketcheson2008highly} & 10&9&8&6&4& --- & ---\\
    Cash--Karp RK5(4)6 \cite{cash1990variable} & 6&5&4&2&1&0& --- \\
    Dormand--Prince RK5(4)7 \cite{prince1981high}& 7&6&5&3&1&0& --- \\
    \bottomrule
  \end{tabular*}
  \caption{Degrees of freedom for the choice of the weights for some explicit methods.} %Generated above============
  \label{table:DOF_exp}
\end{table}

\begin{table}[h!]
\centering   %Generated below============ 
%   \begin{tabular}{|l c|c c c c c c |}
%   \hline
%   Order & &1&2&3&4&5&6 \\
%   \hline Implicit Euler& s=1&0& - & - & - & - & -  \\
%   Lobatto~IIIC4& s=4&3&2&1&0&0&0 \\
%   Radau~IIA3& s=3&2&1&0&0&0& -  \\
%   SDIRK(5,4)& s=5&4&3&1&0& - & -  \\
%   TR-BDF2& s=3&2&1& - & - & - & -  \\
%   Im-Euler 2& s=3&2&1& - & - & - & -  \\
%   Im-Euler 3& s=6&5&4&2& - & - & -  \\
%   Im-Euler 4& s=10&9&8&6&3& - & -  \\
%   \hline
%   \end{tabular}
   \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lr*6c@{}}
    \toprule
    Method & $s$ & \multicolumn{6}{c}{Order $\tilde p$} \\
    & & 1 & 2 & 3 & 4 & 5 & 6 \\
    \midrule
    Implicit Euler& 1&0& --- & --- & --- & --- & ---  \\
    Lobatto~IIIC4 \cite{chipman1971stable} & 4&3&2&1&0&0&0 \\
    Radau~IIA3 \cite{ehle1969pade} & 3&2&1&0&0&0& ---  \\
    SDIRK(5,4) \cite[eq. (6.18)]{hairer_solving_1996}& 5&4&3&1&0& --- & ---  \\
    TR-BDF2 \cite{bank1985transient} & 3&2&1& --- & --- & --- & ---  \\
    Extrapolation Im-Euler 2& 3&2&1& --- & --- & --- & ---  \\
    Extrapolation Im-Euler 3& 6&5&4&2& --- & --- & ---  \\
    Extrapolation Im-Euler 4& 10&9&8&6&3& --- & ---  \\
    \bottomrule
  \end{tabular*}
  \caption{Degrees of freedom for the choice of the weights for some implicit methods.} %Generated above============
  \label{table:DOF_imp}
\end{table}

@@

In Table\,\ref{table:DOF_imp}, the same properties of some implicit
methods are reported.
We can see that the fully implicit methods Lobatto~IIIC4 and Radau~IIA3 require a drastic reduction of the order as expected.
The diagonally implicit SDIRK(5,4) method only requires an order reduction of one to get one degree of freedom for the weights.
The TR-BDF2 method even allows adaptations without reducing the order.  
The Implicit-Euler extrapolation methods also exhibit degrees of freedom without a reduction of the order.
